---
title: "Launching code"
output: html_notebook
---

### Required packages/software:

- `rstan`
- a recent version of *stan* (e.g. 2.26.1 or more recent)


### Input

The input is a csv file containing the biomass of species *j* in plot *i* in cell *i,j*. The first column of the file contains the names of the species. For example:

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
read_csv("../data/cadotte_2013/organized_data/cadotte_2013.csv")
```

In a file with the same name, but ending in "_V.csv" we store the matrix detailing the topology of the tree. Here columns are the species (in the same order as the previous file), rows are ancestors, and a 1 marks the fact that the row/ancestor is a progenitor of the column/species. For example:

```{r, message=FALSE, warning=FALSE}
read_csv("../data/cadotte_2013/organized_data/cadotte_2013_V.csv")
```

## Running the code

For each data set, one can run the code using one of four models (see manuscript). The syntax is as follows:

```{r, message=FALSE, warning=FALSE, eval=FALSE}
source("pipeline.R") # this will take a minute, to compile stan code
run_model(file_obs = "../data/cadotte_2013/organized_data/cadotte_2013.csv", # data, as above
          modelnum = 4, # model number
          seed = 1234, # seed for random number generator
          num_search = 2, # repeat search multiple times and take the best result
          num_optim = 2500, # number of rounds for optimization
          num_rand = 100) # number of randomizations (Yule; another 100 scrambled)
# This code will take about 1 day to run
```

Besides the four models described in the manuscript, calling `11`, `12`, `13` and `14` calls the models using a star/broom tree topology. Running each of these sets the baseline for the likelihoods, as described in the manuscript.

## Organizing results

Because the code takes much time, it is convenient to run several instances in parallel (each starting with a different seed for the random number generator). All the outputs will be stored in the *results* folder.

The code `organize_results.R` will combine multiple runs. It will also remove randomizations that failed (i.e., if likelihood falls below that of the corresponding star tree), and will remove duplicate random trees.


